---
title: "peerfeedback28"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```


## Communication

### 28.2.1.1

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```

Very nicely done - good use of colors with clear labeling.

### 28.2.1.3

```{r}
ggplot(diamonds, aes(x = carat)) + geom_density() +
    labs(title = "The Number of Diamonds Generally decrease with an increase in carat", 
         subtitle = "there are spikes at whole numbers due to human preferance", 
         x = "Count")
```

I like that the title and subtitle specifically tell stories within the data.

### 28.4.4.3

```{r}
presidential %>%
mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_colour_manual(values = c(Republican = "red", Democratic = "blue")) +
  scale_y_continuous(breaks = seq(34, 44, by = 1), labels = presidential$name) +
  labs(x = "year", y = "President", title = "Presidentail Terms") +
  scale_x_date(date_breaks = "4 years", date_labels = "%y")
```

Solid again. Is there a pattern in the data that could have made a fitting title or subtitle?

## Technical Understanding

### 28.3.1

```{r}
mpg2 <- mpg %>% group_by(cyl) %>% mutate(avg_cty = round(mean(cty),1))
ggplot(data = mpg2, aes(cty, hwy)) +
  geom_point() +
  geom_text(aes(label = avg_cty), y = 45, x = 10) +
  facet_grid(~cyl)
```

No worries, the individual labels within facets is definitely a weird thing to hunt down. It involves specifying factor levels within the label; there is a janky implementation of it in my file for an example.

### 28.4.4 (1,2,4)

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(colour = cut), alpha = 1/20) +
  guides(color = guide_legend(title = "Quality",override.aes = list(alpha = 1, shape = 1, size = 4)))
```

Good choice of size and shape in the legend, and alpha in the aesthetics.

## General Notes

Great job on these, Rags. It is clear that you have a very sound understanding of how to create and format plots. One pointer is to make sure the files are saved as the appropriate type (noticed a few files in your repo didn't have an extension) so that they are easily readable/executable.
