---
title: "r4ds_communicate_exercises_27"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## 27.2.1 Exercises

### 1. Create a new notebook using *File > New File > R Notebook*. Read the instructions. Practice running the chunks. Verify that you can modify the code, re-run it, and see modified output.

```{r}
print("Okey dokey")
```

### 2. Create a new R Markdown document with File > New File > R Markdown. Knit it by clicking the appropriate button. Knit it by using the appropriate keyboard short cut. Verify that you can modify the input and see the output update.

Okey dokey.

### 3. Compare and contrast the R notebook and R markdown files you created above. How are the outputs similar? How are they different? How are the inputs similar? How are they different? What happens if you copy the YAML header from one to the other?

For inputs, they can both take R chunks. They can both output results and plots, but the notebook doesn't show the chunks of R code. The notebook doesn't actually run the R code when it is generated; it will only show the output for chunks of R code that you ran manually prior to generating the output doc.

Copying the YAML header of a .Rmd file to a .nb file will result in the .nb file automatically being converted to a .Rmd; it outputs to the document type you specified without issue. Copying the YAML header of a .nb file to a .Rmd file doesn't convert the source file type, and will output to the document type you specified without issue.

### 4. Create one new R Markdown document for each of the three built-in formats: HTML, PDF and Word. Knit each of the three documents. How does the output differ? How does the input differ? (You may need to install LaTeX in order to build the PDF output - RStudio will prompt you if this is necessary.)

They're not that different. You don't really need to change anything in the code other than the `output` argument in the YAML header. PDF is just a heavier file type. And a Word doc looks kind of gross, not sure what type of situation might arise where I'd want to do that.

## 27.3.1 Exercises

### 1. Practice what you've learned by creating a brief CV. The title should be your name, and you should include headings for (at least) education or employment. Each of the sections should include a bulleted list of jobs/degrees. Highlight the year in bold.

#### Calvin Ke

##### Education

* Dank Memer Academy, **2010-2012**
* Vandermeme University, *Summa meme laude* **2013-2016**

##### Employment

* The Memecake Factory, **2016-**

### 2. Using the R Markdown quick reference, figure out how to:

1. Add a footnote.

a [footnote][^1]

2. Add a horizontal rule.

above the fold

---

below the fold

3. Add a block quote.

> be me  
> doing r4ds exercises  
> is this 4chan  

### 3. Copy and paste the contents of `diamond-sizes.Rmd` from https://github.com/hadley/r4ds/tree/master/rmarkdown in to a local R markdown document. Check that you can run it, then add text after the frequency polygon that describes its most striking features.

```{r}
smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only `r nrow(diamonds) - nrow(smaller)` are larger than 2.5 carats. The distribution of the remainder is shown below:

```{r}  
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

The freqpoly shows a higher count around carats with round number values. This indicates that the person logging the data was a jabroni.

### 27.4.7 Exercises

### 1. Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you're writing a report for someone who doesn't know R, and instead of setting `echo = FALSE` on each chunk, set a global option.

Ugh.

#### Variation of Diamond Sizes by Cut, Color, and Clarity

```{r, eval = FALSE}
knitr::opts_chunk$set(
  echo = FALSE
)
```

### 2. Download `diamond-sizes.Rmd` from https://github.com/hadley/r4ds/tree/master/rmarkdown. Add a section that describes the largest 20 diamonds, including a table that displays their most important attributes.

#### 20 Largest Diamonds Summarized

They are pretty dank.

```{r}
d20 <- diamonds %>% 
  arrange(desc(carat)) %>% 
  head(20)

knitr::kable(
  d20,
  caption = "20 dankest diamonds"
)
```

### 3. Modify `diamonds-sizes.Rmd` to use `comma()` to produce nicely formatted output. Also include the percentage of diamonds that are larger than 2.5 carats.

```{r}
comma <- function(x) format(x, digits = 2, big.mark = ",")
d2 <- diamonds %>% 
  mutate(price = comma(price))
```

`r comma(100 * (nrow(diamonds) - nrow(smaller)) / nrow(diamonds))`% of diamonds are larger than 2.5 carats.

### 4. Set up a network of chunks where `d` depends on `c` and `b`, and both `b` and `c` depend on `a.` Have each chunk print `lubridate::now()`, set `cache = TRUE`, then verify your understanding of caching.

```{r a, cache=TRUE}
lubridate::now()
```

```{r b, dependson="a", cache=TRUE}
lubridate::now()
```

```{r c, dependson="a", cache=TRUE}
lubridate::now()
```

```{r d, dependson=c("b", "c"), cache=TRUE}
lubridate::now()
```

Wadu hek. I don't get it. They all update independently regardless of the dependencies.

[^1]: footnote content