---
title: "r4ds_wrangle_exercises_14-15"
output: html_document
---

```{r, message=FALSE, warning=FALSE, results="hide"}
library(tidyverse)
library(stringr)
```

## 14.2.5 Exercises

### 1. In code that doesn't use stringr, you'll often see `paste()` and `paste0()`. What's the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of `NA`?

```{r, eval=FALSE}
?paste
?paste0
stringr::str_c("a", NA)
paste("a", NA)
```

`paste` is like `str_c` except you need to specify a separator (which if you don't the default is a space). `paste0` is `paste` with the separator set to nothing. `str_c` deals with NA values as a "traditional" operation would - trying to concat NA with anything else will result in NA. Whereas paste will literally concat the string "NA" onto whatever else.

### 2. In your own words, describe the difference between the `sep` and `collapse` arguments to `str_c()`.

`sep` operates on individual strings, `collapse` operates on vectors of strings.

### 3. Use `str_length()` and `str_sub()` to extract the middle character from a string. What will you do if the string has an even number of characters?

```{r}
odd <- "horse"
even <- "horseman"

stringr::str_sub(odd, stringr::str_length(odd)/2 + 0.5, stringr::str_length(odd)/2 + 0.5)
stringr::str_sub(even, stringr::str_length(even)/2, stringr::str_length(even)/2 + 1)
```

### 4. What does `str_wrap()` do? When might you want to use it?

```{r, eval=FALSE}
?str_wrap
```

Seems like it would wrap a string onto a new line(s) if it is too long. This could be useful for formatting, if we don't want lines of text to spill over a certain width.

### 5. What does `str_trim()` do? What's the opposite of `str_trim()`?

```{r, eval=FALSE}
?str_trim
```

It removes white space on either side of a string. The opposite is `str_pad()`.

### 6. Write a function that turns (e.g.) a vector `c("a", "b", "c")` into the string `a, b, and c`. Think carefully about what it should do if given a vector of length 0, 1, or 2.

```{r}
vector3 <- c("a", "b", "c")
vector2 <- c("a", "b")
vector1 <- c("a")
vector0 <- c()

str_c(str_c(vector3[-length(vector3)], collapse = ", "), vector3[length(vector3)], sep = ifelse(length(vector3) > 2, ", and ", ", "))
str_c(str_c(vector2[-length(vector2)], collapse = ", "), vector2[length(vector2)], sep = ifelse(length(vector2) > 2, ", and ", ", "))
str_c(str_c(vector1[-length(vector1)], collapse = ", "), vector1[length(vector1)], sep = ifelse(length(vector1) > 2, ", and ", ", "))
str_c(str_c(vector0[-length(vector0)], collapse = ", "), vector0[length(vector0)], sep = ifelse(length(vector0) > 2, ", and ", ", "))
```

## 14.3.1.1 Exercises

### 1. Explain why each of these strings don't match a `\`:

`\` doesn't match because it is interpreted as an escape character rather than the actual character.

`\\` is an escape character in a regexp.

`\\\` is a literal backslash and regexp escape character next to each other.

Ugh, this is getting way to inception-y for me.


### 2. How would you match the sequence `"'\`?

`'"\'\\'`


### 3. What patterns will the regular expression `\..\..\..` match? How would you represent it as a string?

It will match anything that looks like `.%.%.%` where the dots are literal dots and `%` stands for the wild card.

To represent it as a string, we could try `"\\.\\.\\."`.

## 14.3.2.1 Exercises

### 1. How would you match the literal string `"$^$"`?

`"^\\$\\^\\$$"`

### 2. Given the corpus of common words in stringr::words, create regular expressions that find all words that:

### 3. Start with "y".

```{r}
stringr::str_view(stringr::words, "^y.", match = TRUE)
```

### 4. End with "x"
```{r}
stringr::str_view(stringr::words, ".x$", match = TRUE)
```

### 5. Are exactly three letters long.
```{r}
stringr::str_view(stringr::words, "^...$", match = TRUE)
```

### 6. Have seven letters or more.

```{r}
stringr::str_view(stringr::words, ".......", match = TRUE)
```

## 14.3.3.1 Exercises

### 1. Create regular expressions to find all words that:

```{r}
nato <- c("alpha", "bravo", "charlie", "delta", "echo", "foxtrot", "golf", "hotel", "india", "juliet", "kilo", "lima", "mike", "november", "oscar", "papa", "quebec", "romeo", "sierra", "tango", "uniform", "victor", "whiskey", "x-ray", "yankee", "zulu", "asdf")
```

#### 1. Start with a vowel.
```{r}
stringr::str_view(nato, "^(a|e|i|o|u)", match = T)
```

#### 2. That only contain consonants. (Hint: thinking about matching "not"-vowels.)
```{r}
stringr::str_view(nato, ".[^aeiou].", match = T)
```

WADU HEK, why didn't that work...

#### 3. End with `ed`, but not with `eed`.
```{r}
stringr::str_view(c("feed", "fed"), "[^e]ed$")
```

#### 4. End with `ing` or `ise`.
```{r}
stringr::str_view(c("sample", "devising", "demise"), "ing$|ise$")
```

### 2. Empirically verify the rule "i before e except after c".

Y'all already know that shit is a lie.
```{r}
stringr::str_view(stringr::words, "[^c]ei", match = T)
```

Yeah right bruh, only one counterexample is not representative, that is cherrypicking -- I can do it, too: seize, heinous, foreign, neighbor, beige, weird, their.

### 3. Is "q" always followed by a "u"?

```{r}
stringr::str_view(stringr::words, "qu", match = T)
stringr::str_view(stringr::words, "q[^u]", match = T)
```

It would appear so as there are no counterexamples, at least in the `words` set.

### 4. Write a regular expression that matches a word if it's probably written in British English, not American English.

```{r}
stringr::str_view(stringr::words, "our|ise", match = T)
```

Welp, that somewhat worked.

### 5. Create a regular expression that will match telephone numbers as commonly written in your country.

```{r}
sample_numbers <- "555-5555, 1234-5678"
stringr::str_view(sample_numbers, "\\d\\d\\d-\\d\\d\\d\\d")
```

## 14.4.2 Exercises

### 1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple `str_detect()` calls.

#### 1. Find all words that start or end with `x`.
```{r}
stringr::str_view(stringr::words, "(^x)|(x$)", match = T)
xbegin <- stringr::str_detect(stringr::words, "^x")
xend <- stringr::str_detect(stringr::words, "x$")
stringr::words[xbegin | xend]
```

#### 2. Find all words that start with a vowel and end with a consonant.
```{r}
stringr::str_view(stringr::words, "^[aeiou].*[^aeiou]$", match = T)
vbegin <- stringr::str_detect(stringr::words, "^[aeiou]")
cend <- stringr::str_detect(stringr::words, "[^aeiou]$") 
stringr::words[vbegin & cend]
```

#### 3. Are there any words that contain at least one of each different vowel?
```{r}
# not even going to bother trying using str_view for this one
a <- stringr::str_detect(stringr::words, "a")
e <- stringr::str_detect(stringr::words, "e")
i <- stringr::str_detect(stringr::words, "i")
o <- stringr::str_detect(stringr::words, "o")
u <- stringr::str_detect(stringr::words, "u")
stringr::words[a & e & i & o & u]
```

Nope, not in this `words` set.


### 2. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)

```{r}
vwords <- as.tibble(stringr::words) %>% 
  mutate(vcount = stringr::str_count(value, "[aeiou]"))
vwords %>% 
  arrange(desc(vcount)) %>% 
  head()

vwords2 <- vwords %>%
  mutate(vprop = vcount / stringr::str_length(value))
vwords2 %>% 
  arrange(desc(vprop)) %>% 
  head()
```

A lot of words have 5 vowels, so it's a tie. The word "a" has the highest vowel proportion at 100%.

## 14.4.6.1 Exercises

### 1. Split up a string like `"apples, pears, and bananas"` into individual components.

```{r}
stringr::str_split("apples, pears, and bananas", ", (and )?")[[1]]
```

### 2. Why is it better to split up by `boundary("word")` than `" "`?

It's more robust. Splitting by `" "` is also going to capture the space between sentences and include punctuation like commas in the words.

### 3. What does splitting with an empty string (`""`) do? Experiment, and then read the documentation.

```{r}
stringr::str_split("apples, pears, and bananas", "")[[1]]
```

Appears to split every character. According to `?str_split`, this is equivalent to using `boundary("character")`.











